VIPE: Virtualized Intermittently-Powered Executor

### 2017-07-17 paper/background.tex
ASPLOS latex template replaced with a correct one/bib file updated following the style

\subsection{Energy Harvesting: Background}

@@

@@

\TODO{Elaborate on the ``key challenge'' paragraphs near the end of the intro.}

>>>>

\subsection{Energy Harvesting: Background}
\label{sec:background_harvesting}

@@

\subsection{Memory Consistency}
\label{sec:background_consistency}

Naturally, content of a volatile memory (i.e. run-time stack, global and local variables) of the microcontroller and program registers are erased after each power failure. Only what resides in a non-volatile memory, e.g. FRAM, persists after system reboot. A solution to sustain the progress of the program from the last power failure, instead from the beginning (the \emph{Sisyphus} effect~\cite[Sec. 2]{mementos})is a periodic checkpointing of a program state~\cite{mementos,hibernusplusplus,quickrecall,idetic}. The periodicity, and a position of a checkpoint, is the core of the contribution behind each of the proposed checkpoining solution.

\begin{figure}
	\centering
	%\includegraphics[width=0.25\columnwidth]{figures/code_demo_incosistency}
	%	for (byte = 0; byte < nBytes; ++byte){\\
	%	checkpoint();\\
		%data = reflectData(message\[byte\]) $^$ (remainder >> (WIDTH - 8));\\
		%remainder = crcTable[data] $^$ (remainder << 8);}
	\caption{Code demonstrating WAR and volatile memory checkpointing.}
	\label{fig:code_demo_incosistency}
\end{figure}

Fig.~\ref{fig:code_demo_incosistency} demonstrates the problem through example.

@@

%Q1: Does Program Counter change size with task (checkpoint) size in Chain? In other words: bigger Chain task=bigger Program Counter?
%Q2: Does Register File (RF) includes Program Counter (PC)? In other words: is always PC>RF in terms of memory size?

%N_{checkpoints} \times e^{2\times N_{shared}}

\todo{Elaborate on the ``key challenge'' paragraphs near the end of the intro}{Przemek}
\todo{Define the Task and Task based execution model}{Przemek}

### 2017-07-17 paper/intro.tex
ASPLOS latex template replaced with a correct one/bib file updated following the style

\todo{Put exact reference for "tictpl"}{Brandon}
\todo{Check if the list of authors for Alpaca is correct}{Brandon}

@@

With this paper we provide the following list of contributions.
%
\begin{enumerate}
%
\item {\bf Task coalescing mechanism:} \sys's core contribution is the introduction of dynamic task coalescing mechanism. It allows the programmer or a compiler to intuitively decompose the program into small tasks that amortize fixed per-task overheads, yet present no risk of exceeding device energy capacity. As such a decomposition executes, \sys {\em coalesces} dynamically consecutive tasks. Coalescing elides the commits of coalesced tasks by buffering multiple tasks' updates in an FRAM commit buffer. Periodically, as the span of the coalesced task grows, \sys ends coalescing and commits the state of the coalesced task. If a power failure interrupts a sequence of coalesced tasks, \sys adaptively reduces the number of tasks in that sequence that it will
coalesce, committing sooner in future executions. Consequently, \sys's coalescing mechanism allows a program to execute efficiently across a range of
energy buffer sizes, avoiding transition overheads in larger buffers, and
ensuring progress in smaller buffers.
%
\todo{Provide numerical results for this section}{Przemek}
%
\item {\bf Evaluation of data swapping strategies:} We propose and explore two data swapping strategies for intermittently-powered systems based on FRAM/SRAM architecture. The first strategy is {\em demand paging}, which swaps the SRAM page with a new page from FRAM, buffering the swapped-out page until commit. The second strategy is {\em buffered direct access}, which directly reads and writes FRAM relying on dynamic double-buffering to ensure memory is consistent at
commit. 
%
\todo{Once paging is fully implemented--update the paragraph}{Przemek}
\todo{Provide numerical results for this section}{Przemek}
\todo{Find better name for demand paging}{Brandon}
%
\item {\bf Extensive \sys evaluation:} We evaluated \sys on a collection of benchmark programs~\cite[Sec. 5]{chain}, and compare directly to state-of-the art task-based runtime~\cite{chain}, showing that \sys has high performance and can flexible target a variety of platforms without recompiling the code. We further demonstrate the efficiency and versatility of \sys with an end-to-end evaluation of a sensing and data processing application deployed in a lab environment. 
%
\todo{add detailed text about evaluation results}{Przemek}
%
\end{enumerate}

>>>>

\todo{Put reference for "tictpl" somewhere}{Brandon}

@@

With this paper we provide the following list of contributions.
%
\begin{enumerate}
%
\item {\bf Task coalescing mechanism:} \sys's core contribution is the introduction of dynamic task coalescing mechanism. It allows the programmer or a compiler to intuitively decompose the program into small tasks that amortize fixed per-task overheads, yet present no risk of exceeding device energy capacity. As such a decomposition executes, \sys~{\em coalesces dynamically} consecutive tasks. Coalescing elides the commits of coalesced tasks by buffering multiple tasks' updates in an FRAM commit buffer. Periodically, as the span of the coalesced task grows, \sys ends coalescing and commits the state of the coalesced task. If a power failure interrupts a sequence of coalesced tasks, \sys adaptively reduces the number of tasks in that sequence that it will coalesce, committing sooner in future executions. Consequently, \sys's coalescing mechanism allows a program to execute efficiently across a range of energy buffer sizes, avoiding transition overheads in larger buffers, and ensuring progress in smaller buffers.
%
\todo{Provide numerical results for this section}{Przemek}
\todo{Mention automatic task generation}{Przemek}
%
\item {\bf Evaluation of data swapping strategies:} We propose and explore two data swapping strategies for intermittently-powered systems based on FRAM/SRAM architecture. The first strategy is {\em demand paging}, which swaps the SRAM page with a new page from FRAM, buffering the swapped-out page until commit. The second strategy is {\em buffered direct access}, which directly reads and writes FRAM relying on dynamic double-buffering to ensure memory is consistent at
commit. 
%
\todo{Once paging is fully implemented--update the paragraph}{Przemek}
\todo{Provide numerical results for this section}{Przemek}
%
\item {\bf Extensive \sys evaluation:} We evaluated \sys on a collection of benchmark programs~\cite[Sec. 5]{chain}, and compare directly to state-of-the art task-based runtime~\cite{chain}, showing that \sys has high performance and can flexible target a variety of platforms without recompiling the code. We further demonstrate the efficiency and versatility of \sys with an end-to-end evaluation of a sensing and data processing application deployed in a lab environment. 
%
\todo{add detailed text about evaluation results}{Przemek}
%
\end{enumerate}

### 2017-07-17 paper/paper.tex
first paragraph of background ready
Expanding background section
paging related things
change the name, this first name include run-time and our system include run time and compile time components
ASPLOS latex template replaced with a correct one/bib file updated following the style

\documentclass[sigconf,anonymous,review]{acmart}

\usepackage[normalem]{ulem}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{algorithm} % environment for the algorithm code (like figure and table)
\usepackage{algorithmicx} % write the pseudo code
\usepackage[noend]{algpseudocode} % layout of the pseudo code
\usepackage{tikz} % needed for: tikz pics
\usepackage{xcolor}
\usepackage{xspace}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{subfigure}
%\usepackage[inline]{enumitem}

\newcommand{\TODO}[1]{{\bf \em \textcolor{red}{TODO: #1}\xspace}}

\newcommand{\todo}[2]{\textcolor{red}{$\bullet$ \textbf{Task:} #1; \textbf{Responsible:} #2}}

\newcommand{\sys}{VIPRE\xspace}
\newcommand{\sysfull}{Virtualized Intermittently-Powered Runtime Environment\xspace}

\setcopyright{rightsretained}

\acmDOI{http://dx.doi.org/xx.xxxx/xxxxxxx.xxxxxxx}

\acmISBN{978-1-xxxx-xxxx-x/17/11}

\acmConference[ASPLOS]{The 23rd ACM International Conference on Architectural Support for Programming Languages and Operating Systems}{March 24--28, 2017}{Williamsburg, VA, USA}

\acmYear{2017}

\copyrightyear{2017}

\acmPrice{15.00}

@@

\title[VIPRE: Virtualized Intermittently-Powered Runtime Environment]{VIPRE: Virtualized Intermittently-Powered Runtime Environment}

@@

\author{Kiwan Maeng}
%\authornote{}
%\orcid{}
\affiliation{%
	\institution{Carnegie Mellon University}
	\streetaddress{4720 Forbes Avenue}
	\city{Pittsburgh} 
	\state{PA} 
	\postcode{15213}
}
\email{kmaeng@andrew.cmu.edu}

\author{Alexei Colin}
%\authornote{}
%\orcid{}
\affiliation{%
	\institution{Carnegie Mellon University}
	\streetaddress{4720 Forbes Avenue}
	\city{Pittsburgh} 
	\state{PA} 
	\postcode{15213}
}
\email{acolin@andrew.cmu.edu}

\author{Brandon Lucia}
%\authornote{}
%\orcid{}
\affiliation{%
	\institution{Carnegie Mellon University}
	\streetaddress{4720 Forbes Avenue}
	\city{Pittsburgh} 
	\state{PA} 
	\postcode{15213}
}
\email{blucia@andrew.cmu.edu}

\author{Amjad Yousef Majid}
%\authornote{}
%\orcid{}
\affiliation{%
	\institution{Delft University of Technology}
	\streetaddress{Mekelweg 4}
	\city{Delft, The Netherlands} 
	\state{Zuid Holland} 
	\postcode{2628\,CD}
}
\email{a.y.majid@tudelft.nl}

\author{Kas{\i}m Sinan Y{\i}ld{\i}r{\i}m}
%\authornote{}
%\orcid{}
\affiliation{%
	\institution{Delft University of Technology}
	\streetaddress{Mekelweg 4}
	\city{Delft, The Netherlands} 
	\state{Zuid Holland} 
	\postcode{2628\,CD}
}
\email{k.s.yildirim@tudelft.nl}

\author{Przemys{\l}aw Pawe{\l}czak}
%\authornote{}
%\orcid{}
\affiliation{%
	\institution{Delft University of Technology}
	\streetaddress{Mekelweg 4}
	\city{Delft, The Netherlands} 
	\state{Zuid Holland} 
	\postcode{2628\,CD}
}
\email{p.pawelczak@tudelft.nl}

\renewcommand{\shortauthors}{A. Y. Majid et al.}

@@

\begin{CCSXML}
	<ccs2012>
	<concept>
	<concept_id>10010520.10010553.10010562.10010564</concept_id>
	<concept_desc>Computer systems organization~Embedded 
	software</concept_desc>
	<concept_significance>500</concept_significance>
	</concept>
	<concept>
	<concept_id>10010520.10010575.10010578</concept_id>
	<concept_desc>Computer systems organization~Availability</concept_desc>
	<concept_significance>300</concept_significance>
	</concept>
	<concept>
	<concept_id>10011007.10010940.10010941.10010949.10010957.10010688</concept_id>
	<concept_desc>Software and its engineering~Scheduling</concept_desc>
	<concept_significance>300</concept_significance>
	</concept>
	<concept>
	<concept_id>10011007.10010940.10010941.10010949.10010965</concept_id>
	<concept_desc>Software and its engineering~Communications 
	management</concept_desc>
	<concept_significance>300</concept_significance>
	</concept>
	<concept>
	<concept_id>10011007.10010940.10010971.10010564</concept_id>
	<concept_desc>Software and its engineering~Embedded 
	software</concept_desc>
	<concept_significance>300</concept_significance>
	</concept>
	</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Computer systems organization~Embedded software}
\ccsdesc[300]{Computer systems organization~Availability}
\ccsdesc[300]{Software and its engineering~Scheduling}
\ccsdesc[300]{Software and its engineering~Communications management}
\ccsdesc[300]{Software and its engineering~Embedded software}

\keywords{Energy harvesting, transient operation, operating system}

\maketitle

@@

\section{Computation on Intermittent Devices: Background}
\label{sec:background}

@@

\section{Memory Virtualization}
\label{sec:memory_virtulaization}

\begin{figure}
	\centering
	%\includegraphics[width=0.25\columnwidth]{figures/}
	\caption{Memory virtualization architecture.}
	\label{fig:}
\end{figure}

@@

%%%%%

@@

\begin{figure}[t]
    \centering
         \subfigure[The energy cost of FRAM/SRAM \textbf{writes} operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_write.eps} }
         \subfigure[The energy cost of FRAM/SRAM \textbf{read} operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_read.eps}}
    \caption{The energy cost of accessing volatile/non-volatile memory}
    \label{fig:framEnergy}
\end{figure}

@@

\bibliographystyle{abbrvnat}
\bibliography{bib}

>>>>

%\documentclass[sigconf,anonymous,review]{acmart}
\documentclass[pageno]{jpaper}

%replace XXX with the submission number you are given from the ASPLOS submission site.
\newcommand{\asplossubmissionnumber}{XXX}

\usepackage[normalem]{ulem}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{algorithm} % environment for the algorithm code (like figure and table)
\usepackage{algorithmicx} % write the pseudo code
\usepackage[noend]{algpseudocode} % layout of the pseudo code
\usepackage{tikz} % needed for: tikz pics
\usepackage{xcolor}
\usepackage{xspace}
\usepackage{natbib}
%\usepackage{graphicx}
%\usepackage{subfigure}
%\usepackage[inline]{enumitem}

\newcommand{\TODO}[1]{{\bf \em \textcolor{red}{TODO: #1}\xspace}}
\newcommand{\todo}[2]{\textcolor{red}{$\bullet$ \textbf{Task:} #1; \textbf{Responsible:} #2}}
\newcommand{\sys}{VIPE\xspace}
\newcommand{\sysfull}{Virtualized Intermittently-Powered Executor\xspace}

%\setcopyright{rightsretained}

%\acmDOI{http://dx.doi.org/xx.xxxx/xxxxxxx.xxxxxxx}

%\acmISBN{978-1-xxxx-xxxx-x/17/11}

%\acmConference[ASPLOS]{The 23rd ACM International Conference on Architectural Support for Programming Languages and Operating Systems}{March 24--28, 2017}{Williamsburg, VA, USA}

%\acmYear{2017}

%\copyrightyear{2017}

%\acmPrice{15.00}

@@

\title{VIPE: Virtualized Intermittently-Powered Executor}

@@

%\author{Kiwan Maeng}
%\authornote{}
%\orcid{}
%\affiliation{%
%	\institution{Carnegie Mellon University}
%	\streetaddress{4720 Forbes Avenue}
%\city{Pittsburgh} 
%	\state{PA} 
%	\postcode{15213}
%}
%\email{kmaeng@andrew.cmu.edu}

%\author{Alexei Colin}
%\authornote{}
%\orcid{}
%\affiliation{%
%	\institution{Carnegie Mellon University}
%	\streetaddress{4720 Forbes Avenue}
%	\city{Pittsburgh} 
%	\state{PA} 
%	\postcode{15213}
%}
%\email{acolin@andrew.cmu.edu}

%\author{Brandon Lucia}
%\authornote{}
%\orcid{}
%\affiliation{%
%	\institution{Carnegie Mellon University}
%	\streetaddress{4720 Forbes Avenue}
%	\city{Pittsburgh} 
%	\state{PA} 
%	\postcode{15213}
%}
%\email{blucia@andrew.cmu.edu}

%\author{Amjad Yousef Majid}
%\authornote{}
%\orcid{}
%\affiliation{%
%	\institution{Delft University of Technology}
%	\streetaddress{Mekelweg 4}
%	\city{Delft, The Netherlands} 
%	\state{Zuid Holland} 
%	\postcode{2628\,CD}
%}
%\email{a.y.majid@tudelft.nl}

%\author{Kas{\i}m Sinan Y{\i}ld{\i}r{\i}m}
%\authornote{}
%\orcid{}
%\affiliation{%
%	\institution{Delft University of Technology}
%	\streetaddress{Mekelweg 4}
%	\city{Delft, The Netherlands} 
%	\state{Zuid Holland} 
%	\postcode{2628\,CD}
%}
%\email{k.s.yildirim@tudelft.nl}

%\author{Przemys{\l}aw Pawe{\l}czak}
%\authornote{}
%\orcid{}
%\affiliation{%
%	\institution{Delft University of Technology}
%	\streetaddress{Mekelweg 4}
%	\city{Delft, The Netherlands} 
%	\state{Zuid Holland} 
%	\postcode{2628\,CD}
%}
%\email{p.pawelczak@tudelft.nl}

%\renewcommand{\shortauthors}{A. Y. Majid et al.}

\date{}
\maketitle

\thispagestyle{empty}

@@

%\begin{CCSXML}
%	<ccs2012>
%	<concept>
%	<concept_id>10010520.10010553.10010562.10010564</concept_id>
%	<concept_desc>Computer systems organization~Embedded 
%	software</concept_desc>
%	<concept_significance>500</concept_significance>
%	</concept>
%	<concept>
%	<concept_id>10010520.10010575.10010578</concept_id>
%	<concept_desc>Computer systems organization~Availability</concept_desc>
%	<concept_significance>300</concept_significance>
%	</concept>
%	<concept>
%	<concept_id>10011007.10010940.10010941.10010949.10010957.10010688</concept_id>
%	<concept_desc>Software and its engineering~Scheduling</concept_desc>
%	<concept_significance>300</concept_significance>
%	</concept>
%	<concept>
%	<concept_id>10011007.10010940.10010941.10010949.10010965</concept_id>
%	<concept_desc>Software and its engineering~Communications 
%	management</concept_desc>
%	<concept_significance>300</concept_significance>
%	</concept>
%	<concept>
%	<concept_id>10011007.10010940.10010971.10010564</concept_id>
%	<concept_desc>Software and its engineering~Embedded 
%	software</concept_desc>
%	<concept_significance>300</concept_significance>
%	</concept>
%	</ccs2012>
%\end{CCSXML}

%\ccsdesc[500]{Computer systems organization~Embedded software}
%\ccsdesc[300]{Computer systems organization~Availability}
%\ccsdesc[300]{Software and its engineering~Scheduling}
%\ccsdesc[300]{Software and its engineering~Communications management}
%\ccsdesc[300]{Software and its engineering~Embedded software}

%\keywords{Energy harvesting, transient operation, operating system}

@@

\section{Intermittent Computation: Background}
\label{sec:background}

@@

\section{Memory Virtualization}
\label{sec:memory_virtulaization}
\input{paging}
\begin{figure}
	\centering
	%\includegraphics[width=0.25\columnwidth]{figures/}
	\caption{Memory virtualization architecture.}
	\label{fig:}
\end{figure}

@@

%%%%%Old part of the paper - will be successfully moved section by section to a final one%%%%%

\newpage

@@

\begin{figure}[t]
    \centering
         \subfloat[The energy cost of FRAM/SRAM \textbf{writes} operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_write.eps} }
         \subfloat[The energy cost of FRAM/SRAM \textbf{read} operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_read.eps}}
    \caption{The energy cost of accessing volatile/non-volatile memory}
    \label{fig:framEnergy}
\end{figure}

@@

\bibliographystyle{plain}
\bibliography{bib}
